---
title: "SparsePVCM"
author: "Joonho Gong, Luo Xiao, Arnab Maity"
output: rmarkdown::html_document 
vignette: >
  %\VignetteIndexEntry{my-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(SparsePVCM)
```
## Functions

### SparsePVCM
`SparsePVCM` is a function to fit a sparse varying coefficient model for a small number of principal functions and variable selection. 

#### Usage
```{R, eval =F}
### 1. Prepare the yeat real data
data(yeast)

## Time series response variable
y_sample <- c( t(yeast$y) )
n <- length(y_sample)

## Standardize and expand predictor matrix for each response value
X_std <- scale( yeast$x )
n_gene <- dim(X_std)[1]
p <- dim(X_std)[2]

X_sample <- matrix(NA, n, p )
for(i in 1:n_gene){
  temp <-  matrix( rep(X_std[i,], 18), 18, p, byrow = T)
  X_sample[ ( 1 + 18 * (i-1)):(18 * i), ] <- temp
}
X_sample <- cbind( rep(1, n), X_sample)
colnames(X_sample) <- c("intercept", colnames(yeast$x))

## 18 equally spaced time points (0 - 119 minutes)
U_sample <- rep(seq(0,119,length.out= 18), n_gene )


### 2. Fit SPVCM
start.time <- Sys.time()
SparsePVCM_fit <-  SparsePVCM(y_sample, X_sample, U_sample,num_tun_1 = 20, num_tun_2 = 20)
end.time <- Sys.time()
end.time - start.time

start.time <- Sys.time()
SparsePVCM_fit <-  SparsePVCM(y_sample, X_sample, U_sample, num_tun_1 = 20, num_tun_2 = 20)
end.time <- Sys.time()
end.time - start.time

cat(length(SparsePVCM_fit$X_selected), " predictors are selected with rank ", SparsePVCM_fit$rank,"\n")
print(SparsePVCM_fit$X_selected)

### 3. Table 3 in-sample result
# location of the 21 experimentally confirmed TFs with respect to the columns of yeast$x
locat_true <- c(T, T, F, F, F, F, F, F, T, F, T, F, F, F, F, F, F, F,F, F,T, T, F, F, F, F, T, T, T, F, F,
                F, F, F, F, F, F, F, F, F, F ,F, F, F, F, T, F, F, F, F, T, T, T, F, F, F, F, F, F, F, T,
                F, F, F, F, F, F, F, F, T, F, F,
                F, F, F, F, F, F, F, F, F, F, F, T, F, F, F, T, T, F, F, F, T, T, T, F, F, F, F ,F, F, F,F, F, F,F)

Acc_pvcm <- sum( SparsePVCM_fit$nonzero_index[-1] == locat_true ) /106 # Except intercept
MSE_pvcm <- sum(( y_sample - SparsePVCM_fit$y_fitted)^2) / n
Rsqrd_pvcm <- 1- sum(( y_sample - SparsePVCM_fit$y_fitted)^2) / sum(( y_sample - mean(y_sample)  )^2)

TP <- sum(SparsePVCM_fit$nonzero_index[-1] * locat_true) 
FN <- sum( (!(SparsePVCM_fit$nonzero_index[-1])) * locat_true ) 
TN <- sum( (!(SparsePVCM_fit$nonzero_index[-1])) * (!(locat_true)) )
FP <- sum( SparsePVCM_fit$nonzero_index[-1] * (!(locat_true)) ) 

TPR <- TP/ ( TP + FN )
FPR <- FP / ( FP + TN )
print(c(TP, FN, FP, TN))
print(c(Rsqrd_pvcm, MSE_pvcm, Acc_pvcm, TPR, FPR))
```


#### Argument

**y_sample:** The vector of a response variable.

**X_sample:** The design  matrix of predictors.

**U_sample:** The vector of an index variable.

**num_core** The number of cores for parallel computing. If it is not given, the total number of cores minus 1 will be used.

**K:** The number of cubic B-spline basis functions. Default is 10.

**num_tau_1:** The number of nonzero tuning parameter candidates for nuclear norm. Default is 10.

**num_tau_2:** The number of nonzero tuning parameter candidates for group lasso. Default is 10.

**max_iter:** The maximum number of iterations in DYS. Default is 100.

**tol:** Tolerance value used as a stopping criterion. Default is 1e-5.


#### Value
**Theta:**  The estimated parameter matrix \(\hat{\Theta}_{2}\).

**rank:**  The estimated rank of \(\hat{\Theta}_{2}\). 

**X_selected:**  The names of predictors whose have nonzero coefficient functions. 

**nonzero_index:** The logical vector to indicate the predictors with nonzero coefficient functions.

**U:**  The left singular vectors of \(\hat{\Theta}_{2}=UDV^{T}\).

**D:**  The diagonal matrix of singular values of \(\hat{\Theta}_{2}=UDV^{T}\).

**V:**  The right singular vectors of \(\hat{\Theta}_{2}=UDV^{T}\).


### predict_SparsePVCM
`predict_SparsePVCM` is a function to evaluate the estimated functions at the given values of index variable and to calculate the fitted values of the response if new predictor values are given.

#### Usage
```{R, eval =F }
## 4. Coefficient functions
# Figure 2
# evaluate estimated beta and gamma functions at u_new
u_new <- seq(0,119, length.out= 18 )
# For more smooth function, consider
# u_new <- seq(0,119, length.out= 50 )

function_pred <- predict_SparsePVCM(fit = SparsePVCM_fit, U_new = u_new, X_new = NULL)


par(mfrow=c(4,4))
for(i in (1:p)[SparsePVCM_fit$nonzero_index & c(T, locat_true)] ){
  if(i == 1){
    next ## Ignore the intercept function
  }
  plot(
    u_new,
    function_pred$beta_mat[i, ],
    lwd = 2,
    lty = 4,
    col = "red",
    ylim= c(-0.21, 0.37),
    ylab=expression(hat(beta)[j]), xlab="", main=paste(c("Intercept",colnames(yeast$x))[i]),
    type="l",
    xaxt = "n", yaxt = "n",
    cex=1.2,
    cex.axis=1.4,
    cex.lab=1.4
  )
  mtext("Minutes", side = 1, line = 2.5, cex = 0.9)
  abline(h=0, lty=3)
  
  # X-axis
  axis(1, at = seq(0,120, by = 20), cex.axis=1.3)

  # Y-axis
  axis(2, at = seq(-0.2,0.3,by=0.1), labels = c(-0.2, "", 0, "", 0.2,""), cex.axis=1.3)
  
}


```

#### Argument
**fit:** A list that is output of the `SparsePVCM` function.

**U_new:** A numeric vector of index variable values to evaluate estimated functions.

**X_new:** A vector of new predictor values to get fitted values of the response. If not specified with NULL (default), only coefficient functions are evaluated.

#### Value
**y_new:**  The fitted values for the given index variable and predictors.

**beta_mat:**  The estimated coefficient functions \(\hat{\beta}(u)\) at the given index variable. 

**gamma_mat:**  The estimated principal functions \(\hat{\gamma}(u)\) at the given index variable. 


